---
title: "Illinois Median Income"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
editor: 
  markdown: 
    wrap: sentence
---

[Home page](https://nmisetic.github.io/mywebsite/)

## Preparation

First, we must download the necessary packages so that RStudio can perform these functions properly:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(sf)
library(tidycensus)
library(htmltools)
library(janitor)
library(here)
library(mapview)
library(leafsync)
library(leaflet.extras2)
options(tigris_class = "sf")

```

Next, we are going to import census data.
For our purposes, we will use data from 2021.
We will also generate a table to demonstrate all possible conditions

```{r}
myapikey <- "7ec164ce2e6d7eff05a3da72fa85dec41773e398"
census_api_key(myapikey, overwrite = TRUE)

v20 <- load_variables(2021, "acs5", cache=TRUE)
```

## Finding the Data

With all the information properly imported, we can begin collecting our data.
Using the code below, we can sort the data into counties.
Next we use the census key to only highlight median income.
Finally, we must filter for the state of Illinois itself

```{r}
il_median_income <- get_acs(geography = 'county', 
                         variables = c(median_income = 'B19013_001'), 
                         state = 'IL',
                         output = "wide",
                       geometry = TRUE)
il_median_income
```

Next we will remove some unnecessary columns to make the data easier to look at

```{r}
il_median_income <- il_median_income %>%
  select(-ends_with("M"))

colnames(il_median_income) <- sub("E$", "", colnames(il_median_income))

il_median_income
```

## Making the Map

Now, we can generate a map of the state with the following code:

```{r}
mapview(il_median_income, zcol = "median_income")
```

As we can see, only Chicago and Monroe County residents make above the median income
